---
title: "Posterior predictive check for bulk RNA sequencing data"
output: github_document
---

```{r echo=FALSE}
knitr::opts_chunk$set( fig.path = "man/figures/")
```

The input data set is a tidy representation of a differential gene transcript abundance analysis

```{r echo=FALSE, include=FALSE}
library(tidyverse)
library(ppcseq)
library(dplyr)
library(magrittr)
```

To install:

Before install, for linux systems, in order to exploit multi-threading, from R write (without changing anything of the code):

```{r eval=FALSE}
fileConn<-file("~/.R/Makevars")
writeLines(c( "CXX14FLAGS += -O3","CXX14FLAGS += -DSTAN_THREADS", "CXX14FLAGS += -pthread"), fileConn)
close(fileConn)
```

Then, install with

```{r eval=FALSE}
devtools::install_github("stemangiola/ppcseq")
```

You can get the test dataset with

```{r}
data("counts")
counts 
```

You can identify anrtefactual calls from your differential transcribt anundance analysis, due to outliers.

```{r warning=FALSE, message=FALSE,results='hide'}
# Import libraries


counts.ppc = 
	counts %>%
	mutate(is_significant = FDR < 0.01) %>%
	identify_outliers(
		formula = ~ Label,
		.sample = sample, 
		.transcript = symbol,
		.abundance = value,
		.significance = PValue,
		.do_check = is_significant,
		percent_false_positive_genes = 5
	)
```

The new posterior predictive check has been added to the original data frame

```{r }
counts.ppc 
```

The new data frame contains plots for each gene

We can visualise the top five differentially transcribed genes

```{r }
counts.ppc_plots = 
	counts.ppc %>% 
	plot_credible_intervals() 
```

```{r}
counts.ppc_plots %>%
	pull(plot) %>% 
	.[seq_len(length.out=2)]
```
