---
title: "Approximation bias correction"
author: "Mangiola Stefano"
date: "11/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(viridis)
library(tidyverse)

my_theme =
	theme_bw() +
	theme(
		panel.border = element_blank(),
		axis.line = element_line(),
		panel.grid.major = element_line(size = 0.2),
		panel.grid.minor = element_line(size = 0.1),
		text = element_text(size=12),
		legend.position="bottom",
		aspect.ratio=1,
		strip.background = element_blank(),
		axis.title.x  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10)),
		axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
	)
```

## R Markdown

```{r load data, cache=TRUE}
load("dev/inference_before_correction.rda")
```

## Absolute error

The analysis approximation method is introducing some biases. It underestimates the upper quantiles and overestimates the lower quantiles

```{r asolute error, echo=FALSE, cache=TRUE}
inference_before_correction %>%
	filter(Difference != 0) %>%
	filter(p==2 & title=="Approximate posterior analysis") %>%
	left_join(
		ppcSeq::counts %>% group_by( symbol) %>% summarise(value = value %>% `+` (1) %>% log %>% mean, PValue = PValue %>% log %>% mean)
	) %>%
	#sample_frac(0.05) %>%
	{
		my_df = (.)
		ggplot(my_df, aes(x = interaction(sample, symbol), y=Difference, group=sign, sample = sample, symbol=symbol, color = value)) +

			geom_point(alpha=0.5, size=0.1) +
			scale_color_viridis(option="magma") +

			facet_grid(Comparison ~ adj_prob_theshold_2, scales = "free_y") +
			theme(axis.line = element_line(),
						axis.title.x=element_blank(),
						axis.text.x=element_blank(),
						axis.ticks.x=element_blank(),
						#legend.position = "none",
						text = element_text(size=12),
						strip.background = element_blank(),
						axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
			)
	}
```

## Error relative to estimate value

There is a clear correlation with mu, for lower quantiles it seems pretty linear, for upper quantiles it could be exponential

```{r mu error, echo=FALSE, cache=TRUE}
inference_before_correction %>%
	filter(Difference != 0) %>%
	filter(p==2 & title=="Approximate posterior analysis") %>%
	left_join(
		ppcSeq::counts %>% group_by( symbol) %>% summarise(value = value %>% `+` (1) %>% log %>% mean, PValue = PValue %>% log %>% mean)
	) %>%
	ggplot( aes(x = mean_2.x, y=Difference, group=sign, sample = sample, symbol=symbol)) +

	geom_point(alpha=0.5, size=0.1) 	+		facet_grid(Comparison ~ adj_prob_theshold_2, scales = "free_y") +
	my_theme +
	theme(axis.line = element_line(),
				legend.position = "none",
				text = element_text(size=12),
				strip.background = element_blank(),
				axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
	)
```

## Error relative to overdispersion

The sutuation is much less clear with overdispersion. However some sort of association is present.

```{r sigma error, echo=FALSE, cache=TRUE}
inference_before_correction %>%
	filter(Difference != 0) %>%
	filter(p==2 & title=="Approximate posterior analysis") %>%

	ggplot( aes(x = exp(sigma_raw), y=(Difference), group=sign, sample = sample, symbol=symbol)) +

	geom_point(alpha=0.5, size=0.1) 	+		facet_grid(Comparison ~ adj_prob_theshold_2, scales = "free_y") +
	my_theme +
	theme(axis.line = element_line(),
				legend.position = "none",
				text = element_text(size=12),
				strip.background = element_blank(),
				axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
	)
```

## UPPER quantiles - Relation between log(mu) 1/exp(sigma) and log(error)

```{r relation up, cache=TRUE}

inference_before_correction%>%
	filter(Difference != 0) %>%
	filter(p==2 & title=="Approximate posterior analysis") %>%

	filter(Comparison == ".upper_diff") %>%
	sample_frac(0.5) %>%
	ggplot( aes(x = intercept, y=sigma_raw, group=sign, sample = sample, symbol=symbol, color = Difference)) +

	geom_jitter() 	+		facet_grid(Comparison ~ adj_prob_theshold_2, scales = "free_y") +
	scale_color_viridis(option="magma", trans = "log") +
	my_theme +
	theme(axis.line = element_line(),
				#legend.position = "none",
				text = element_text(size=12),
				strip.background = element_blank(),
				axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
	)

```

## lower quantiles - Relation between log(mu) 1/exp(sigma) and log(error)

```{r relation down, cache=TRUE}

inference_before_correction%>%
	filter(Difference != 0) %>%
	filter(p==2 & title=="Approximate posterior analysis") %>%

	filter(Comparison == ".lower_diff") %>%
	sample_frac(0.5) %>%
	ggplot( aes(x = intercept, y=sigma_raw, group=sign, sample = sample, symbol=symbol, color = -Difference)) +

	geom_jitter() 	+		facet_grid(Comparison ~ adj_prob_theshold_2, scales = "free_y") +
	scale_color_viridis(option="magma", trans = "log") +
	my_theme +
	theme(axis.line = element_line(),
				#legend.position = "none",
				text = element_text(size=12),
				strip.background = element_blank(),
				axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
	)

```


## Quantiles affects upper bias exponentially

```{r quantiles, cache=TRUE}

my_medians = inference_before_correction %>%
            filter(Difference != 0) %>%
            filter(p==2 & title=="Approximate posterior analysis") %>%
            
            filter(Comparison == ".upper_diff") %>%
            mutate(Difference = (log(Difference)-intercept)) %>%
            group_by(adj_prob_theshold_2) %>% 
            summarise(Difference = Difference %>% median(na.rm = T)) %>% 
            ungroup()

inference_before_correction %>%
    filter(Difference != 0) %>%
    filter(p==2 & title=="Approximate posterior analysis") %>%
    
    filter(Comparison == ".upper_diff") %>%
    sample_frac(0.5) %>%

    ggplot( aes(x = log(adj_prob_theshold_2), y=log(Difference)-intercept, color = intercept)) +
    
    geom_jitter(alpha = 0.2) 	+	
	# geom_point(
	# 	data = my_medians ,
	# 	aes(x = log(adj_prob_theshold_2), y=(Difference))
	# ) +
	geom_smooth(method  = "lm", color = "black") +
	scale_color_viridis(option="magma", trans = "log") +
	my_theme +
    theme(axis.line = element_line(),
          #legend.position = "none",
          text = element_text(size=12),
          strip.background = element_blank(),
          axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
    )

inference_before_correction %>%
    filter(Difference != 0) %>%
    filter(p==2 & title=="Approximate posterior analysis") %>%
    
    filter(Comparison == ".upper_diff") %>%
    sample_frac(0.5) %>%
    mutate(Difference = (log(Difference)-intercept)) %>%
    group_by(adj_prob_theshold_2) %>% summarise(Difference = Difference %>% median(na.rm = T)) %>% ungroup() %>%
    
    ggplot( aes(x = log(adj_prob_theshold_2), y=(Difference))) +
    
    geom_jitter() 	+		
    #scale_color_gradient( trans = "log") +
	my_theme +
    theme(axis.line = element_line(),
          #legend.position = "none",
          text = element_text(size=12),
          strip.background = element_blank(),
          axis.title.y  = element_text(margin = margin(t = 10, r = 10, b = 10, l = 10))
    )
```


## Build linear models for errors

Upper quantiles

```{r model up, cache=TRUE}

# Build linear model - UPPER diff
inference_before_correction%>%
	filter(Difference != 0) %>%
	filter(p==2 & title=="Approximate posterior analysis") %>%


	filter(Comparison == ".upper_diff") %>%
	group_by(symbol, intercept, sigma_raw, adj_prob_theshold_2) %>%
	summarise(Difference = Difference %>% median) %>%
	ungroup() %>%
	lm(log(Difference) ~ intercept + sigma_raw + log(adj_prob_theshold_2), data = .) %>%
	{
		lm_approx_bias_upper = (.)
		save(lm_approx_bias_upper, file = "dev/lm_approx_bias_upper.rda")
		(.) %>% summary
	}

```

Lower quantiles (here the difference is negative)

```{r model down, cache=TRUE}

# Build linear model - LOWER diff
inference_before_correction%>%
	filter(Difference != 0) %>%
	filter(p==2 & title=="Approximate posterior analysis") %>%

	filter(Comparison == ".lower_diff") %>%
	group_by(symbol, intercept, sigma_raw, adj_prob_theshold_2) %>%
	summarise(Difference = Difference %>% median) %>%
	ungroup() %>%
	lm(log(-Difference) ~ intercept + sigma_raw + log(adj_prob_theshold_2), data = .) %>%
	{
		lm_approx_bias_lower = (.)
		save(lm_approx_bias_lower, file = "dev/lm_approx_bias_lower.rda")
		(.) %>% summary
	}

```



