---
title: "Overview of the tidySingleCellExperiment package"
author: "Stefano Mangiola"
date: "`r Sys.Date()`"
package: tidySingleCellExperiment
output:
  BiocStyle::html_document:
    toc_float: true
bibliography: tidySingleCellExperiment.bib
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Overview of the tidySingleCellExperiment package}
  %\usepackage[UTF-8]{inputenc}
---

```{r, echo=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
    cache=TRUE, warning=FALSE,
    message=FALSE, cache.lazy=FALSE
)
```

The input data set is a tidy representation of a differential gene transcript abundance analysis

```{r echo=FALSE, include=FALSE}
library(dplyr)
library(ppcseq)
library(dplyr)
library(magrittr)
```

To install:

Before install, for linux systems, in order to exploit multi-threading, from R write:

```{r eval=FALSE}
fileConn<-file("~/.R/Makevars")
writeLines(c( "CXX14FLAGS += -O3","CXX14FLAGS += -DSTAN_THREADS", "CXX14FLAGS += -pthread"), fileConn)
close(fileConn)
```

Then, install with

```{r eval=FALSE}
devtools::install_github("stemangiola/ppcseq")
```

You can get the test dataset with

```{r}
ppcseq::counts 
```

You can identify anrtefactual calls from your differential transcribt anundance analysis, due to outliers.

```{r warning=FALSE, message=FALSE,results='hide'}
# Import libraries


counts.ppc = 
	ppcseq::counts %>%
	mutate(is_significant = FDR < 0.01) %>%
	identify_outliers(
		formula = ~ Label,
		.sample = sample, 
		.transcript = symbol,
		.abundance = value,
		.significance = PValue,
		.do_check = is_significant,
		percent_false_positive_genes = 5
	)
```

The new posterior predictive check has been added to the original data frame

```{r }
counts.ppc 
```

The new data frame contains plots for each gene

We can visualise the top five differentially transcribed genes

```{r }
counts.ppc_plots = 
	counts.ppc %>% 
	plot_credible_intervals() 
```

```{r}
counts.ppc_plots %>%
	pull(plot) %>% 
	.[1:2]
```

```{r softwareinfo}
sessionInfo()
```
